---
title: "Lab 8 Assignment"
author: "Jerry Xing"
date: "Oct 30, 2025"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(mosaic)
```

## Experiment 1: Waiting for a Bus

```{r}
X.min <- 0
X.max <- 20
```

#### Q1a

We can sample $m=100$ values of $X$ using `runif`. A histogram of the data follows.

```{r}
# code to generate the X values
hist.X.vals <- function(vals){
  histogram(
    ~vals,
    type = "density",
    right = FALSE,
    breaks = seq(0, 20, by = 1),
    main = paste("Histogram of X for m =", length(vals)),
    xlab = "X values"
  )
}

#now make a histogram
X.vals.100 <- runif(n = 100, min = 0, max = 20)
hist.X.vals(X.vals.100)

X.vals.100000 <- runif(n = 10^5, min = 0, max = 20)
hist.X.vals(X.vals.100000)
```

Using the $m=100$ simulated values, the probability that $X<6$ is:

```{r}
p.hat.100 <- mean(X.vals.100 < 6)
p.hat.100
```

Now do it again with $m=10^5$.

```{r}
p.hat.100000 <- mean(X.vals.100000 < 6)
p.hat.100000
```

#### Q1b

When m becomes larger, the distribution looks more uniform, more consistent with the name "uniform distribution"

#### Q1c

Assuming $X$ is uniformly distributed on the interval $[0, 20]$, the probability that $X < 6$ is:

$X \sim Unif[a,b]$

$P(X<6) = \frac{6-0}{20-0} = \frac{6}{20} = 0.3$


## Experiment 2: Still Waiting (Exponential)

```{r}
beta <- 20 #mean waiting time
lambda <- 1 / beta
```

#### Q2a

```{r}
pexp(6, lambda)
```
#### Q2b

```{r}
# P(X < 10)
p.less.10 <- pexp(10, lambda)

# P(X > 15)
p.greater.15 <- 1 - pexp(15, lambda)

# P(5 < X < 10)
p.between.5.10 <- pexp(10, lambda) - pexp(5, lambda)

c("P(X<10)" = p.less.10,
  "P(X>15)" = p.greater.15,
  "P(5<X<10)" = p.between.5.10)
```

#### Q2c

```{r}
median_wait <- qexp(0.5, lambda)
median_wait

pexp(15, 1/20)
```

#### Q2d

```{r}
Q1 <- qexp(0.25, lambda)
Q3 <- qexp(0.75, lambda)
IQR <- Q3 - Q1

c(Q1 = Q1, Q3 = Q3, IQR = IQR)
```

#### Q2e

The distribution of X is positively skewed.

```{r}
mu  <- beta
sdX <- beta
3 * (mu - median_wait) / sdX
```

#### Q2f

```{r}
hist.exp <- function(vals, m){
  histogram(
    ~ vals,
    type  = "density",
    right = FALSE,
    main  = paste("Histogram of Exponential(β=20), m =", m),
    xlab  = "Waiting time X (minutes)"
  )
}

X.exp.100 <- rexp(n = 100, lambda)
hist.exp(X.exp.100, 100)

X.exp.100000 <- rexp(n = 10^5, lambda)
hist.exp(X.exp.100000, 100000)
```

As m increases from 100 to 100000, the histogram becomes smoother and aligns more closely with the exponential shape.

#### Q2g

```{r}
mean(rexp(n = 10^5, lambda) < 6)
```

#### Q2h

$X \sim Exp(\lambda = \frac{1}{\beta})$

$P(X<6) = 1 - e^{- \frac{6}{20}} \approx 0.2592$

## Experiment 3: Quality Control (Normal Distribution)

```{r}
mu <- 9.01
sig <- 0.05
```

#### Q3a

```{r}
p.less.9.03 <- pnorm(9.03, mu, sig)
p.greater.9.02 <- 1 - pnorm(9.02, mu, sig)
p.between.8.9.9.1 <- pnorm(9.1,mu, sd = sig) - pnorm(8.9,  mu, sig)

c("P(X<9.03)"=p.less.9.03,
  "P(X>9.02)" = p.greater.9.02,
  "P(8.9<X<9.1)" = p.between.8.9.9.1)
```

#### Q3b

```{r}
q.95 <- qnorm(0.95, mu, sig)
q.5 <- qnorm(0.05, mu, sig)
q.25 <- qnorm(0.25, mu, sig)

c("Q95" = q.95, "Q5" = q.5, "Q25" = q.25)
```

#### Q3c

```{r}
X.sim.100 <- rnorm(100, mu, sig)
mean(X.sim.100 > 8.9 & X.sim.100 < 9.1)

X.sim.100000 <- rnorm(10^5, mu, sig)
mean(X.sim.100000 > 8.9 & X.sim.100000 < 9.1)
```

#### Q3d

$P(8.9 < X < 9.1) = P(\frac{8.9 - \mu}{\sigma} < Z < \frac{9.1 - \mu}{\sigma}) = P(-2.2 < Z < 1.8) = \Phi(1.8) - \Phi(-2.2) = \Phi(1.8) - (1 - \Phi(2.2))$

$\Phi(1.8)\approx 0.9641, \Phi(2.2)\approx 0.9861 \rightarrow \Phi(-2.2)=1-0.9861=0.0139$

$\Rightarrow\ P(8.9<X<9.1)\approx 0.9641-0.0139=0.9502$
